%\documentclass[12pt, twocolumn]{article}
\documentclass[12pt]{article}
%\usepackage{fullpage}           % makes all margins 1 inch?
\topmargin=-1.0cm
\textheight=23cm
\evensidemargin=-1.0cm
\oddsidemargin=-1.0cm
\textwidth=19cm
\setcounter{secnumdepth}{-1}     % suppress numbering of sections
\usepackage{amsmath}
\usepackage{amssymb}             % for mathbb
%\usepackage[latin1]{inputenc}    % for ä,ö,ü - doesn't work, we use ae,oe,ue
\usepackage{hyperref}

\begin{document}
	
% formatting:
\parindent=0in
\parskip=0pt
\pagenumbering{roman}
		
% main text
\pagenumbering{arabic} \setcounter{page}{1}

\title{Random Math Formulas}
\maketitle

\section{}
This document contains miscellaneous math formulas gathered from various sources that may or may not become useful at some point. A special focus is on recipe-like formulas that are suitable for being translated into code to become part of the library someday.

\tableofcontents





\section{Linear Algebra}

%\section{Multilinear Algebra}
 % covectors/k-forms, tensors, ..

\section{Calculus}
Calculus is the study of continuous functions and provides the tools for computing slopes of tangents, areas under curves and identifying important features such as extrema and inflection points. Functions can also be unknowns which in the context of calculus can be determined by differential- or integral equations. This is especially important in the physical sciences. It also provides tools to approximate a given function by simpler functions that are well understood, most notably by Taylor- and Fourier series. This is important in numerical computations.
\subsection{Differentiation}
\subsection{Integration} 
 \include{Integration} 
\subsection{Differential Equations}
%\subsection{Limits}
%\subsection{Series}
%\subsection{Integral Transforms} % Fourier- and Laplace trafo
%\subsection{Integral Equations}
%ToDo: finite differences, summation calculus, difference equations, z-trafo
% see https://www.youtube.com/watch?v=4AuV93LOPcE

%\section{Multivariable Calculus}
% scalar fields, vector fields, partial differentiation (grad, Jacobian, Hessian, Laplacian, divergence, curl), integral theorems (Green, Gauss, Stokes), differential forms, generalized Stokes theorem, Tensor fields

%\section{Approximation}
%\paragraph{Trigonometric Interpolation} We have a bunch of datapoints $x_k, y_k$ and want to find 



\begin{thebibliography}{10}
\bibitem{MatheRezepte}
Christian Karpfinger. \newblock {\em Hoehere Mathematik in Rezepten, 3. Auflage}.
\end{thebibliography}	

\bibliography{Bibilography} 
\bibliographystyle{plain}  
	
\end{document}