\section{Universal Algebra}
In \emph{universal algebra}, we take a birds eye view over the ideas that are common to all the specific algebraic structures (such as groups, rings, fields, etc.) that we have seen so far and many more. These common themes are: We are dealing with structures that involve \emph{sets}, \emph{operations} and \emph{axioms}. The sets constitute the backdrop, the operations define the ways in which we can combine set elements to obtain new set elements and the axioms tell us how we expect these operations to behave. For example, in the group $\mathcal{G} = (\mathbb{Z},+)$ of integers under addition, our set is the set of integers, we have one binary operation that we call addition and we have the group axioms that tell us, that this so called addition should be associative, there should be a neutral element and so on. ...TBC...


\subsection{Algebraic Structures}
An \emph{algebraic structure} (aka an \emph{algebraic system} or just an \emph{algebra}\footnote{The term "algebra" may lead to confusion, though because it may mean different things in other contexts.}) consists of the following ingredients: ...TBC...

% Ingredients: 
% (1) A set called the carrier or domain or universe
%     https://mathworld.wolfram.com/UniversalAlgebra.html
% (2) A collection of operations, i.e. functions that take some number of inputs from the set
%     and produce a new element from the set
% (3) A collection of axioms

% ToDo: Mention that the idea of an algebraic structure can be further generalized to allow multiple
% carrier sets, relations instead of (or in addition to) functions, partial functions, infinitary 
% operations, topological structures, See:
% https://de.wikipedia.org/wiki/Universelle_Algebra#Verallgemeinerungen
% ..so, the diea of an algebraic structure un universal algebra it's not so universal after all.


\subsection{Algebraic Specification}
% Closely related to universal algebra as a subfield of mathematics is the subfield of theoretical computer science called algebraic specification or formal specification.

% Differences: in formal specification we make a distinction between the abstract syntactic world defined by a signature and the concrete semantic world defined by an algebra that implements a given signature. Also, in formal specification, we don't deal with axioms (at least not primarily) whereas in universal algebra, the axioms are the focal point (verify!)


\subsubsection{Data Structures}
% ToDo: See pg 112 ff in "Mathematisch-strukturelle Grundlangen der Informatik"

\subsubsection{Signatures}

% A \emph{signature} names abstract symbols for sorts (i.e. data types) and operations. Signatureas are often denoted by a greek uppercase sigma, i.e. by the symbol $\Sigma$. This $\Sigma$ is then used as a name or stand in for the signature just like the letter $f$ is often used as a name for a function.

\subsubsection{Algebras}

% An \emph{algebra} to a signature $\Sigma$ is called a $\Sigma$-Algebra\footnote{Do not confuse this term with the same term used in measure theory! There, $\Sigma$-Algebra refers to an entirely different concept. It's also different from the usage of the term "algebra" in the context of vector spaces where the term means yet a different thing, which is also distinguished from the usage of the term "algebra" as a subfield of math. There are at least 4 different meanings of the word "algebra" in math. It's a total mess!} and it assigns actual concrete sets the abstract sort symbols in $\Sigma$ and assigns concrete functions (aka maps) to the abstract operation symbols in $\Sigma$.

% https://en.wikipedia.org/wiki/Algebraic_specification
% https://de.wikipedia.org/wiki/Spezifikation#Algebraische_Spezifikation

% https://www.informatik.uni-bremen.de/theorie/teach/algspec/2008/Skript/
% https://www.informatik.uni-bremen.de/theorie/teach/algspec/2008/Skript/Skript.pdf
% https://www.informatik.uni-bremen.de/theorie/teach/algspec/2008/Folien/Foliensammlung.pdf

\begin{comment}

https://en.wikipedia.org/wiki/Universal_algebra
https://en.wikipedia.org/wiki/Algebraic_structure
https://en.wikipedia.org/wiki/Variety_(universal_algebra)
https://en.wikipedia.org/wiki/Operad


https://mathworld.wolfram.com/UniversalAlgebra.html
https://de.wikipedia.org/wiki/Universelle_Algebra


According to wikipedia, an algebraic structure is defined via a set and a collection of operations
on the set. Wouldn't it be more general to have a collection of sets and operations between elements
of those sets. Like, for example, the sets could be Nat and Bool and we could have operations like
the unary isEven: Nat -> Bool or something like the binary isDivisibleBy: Nat x Nat -> Bool? Could
functions also have multiple outputs, like for example: quotientAndRemainder Nat x Nat  ->  Nat x Nat
or cartesianToPolar: Real x Real  ->  Real x Real


See:
https://en.wikipedia.org/wiki/Universal_algebra#Generalizations
but that also doesn't talk about collections of sets.

Aha! The more general structure seems to be a heterogeneous algebra:

https://de.wikipedia.org/wiki/Universelle_Algebra#Verallgemeinerungen
https://de.wikipedia.org/wiki/Heterogene_Algebra
https://www.sciencedirect.com/science/article/pii/S002198007080014X


\end{comment}