\section{Vector Spaces}
For a \emph{vector space}, we need two ingredients: \emph{scalars} and \emph{vectors}. The scalars are elements of a field $\mathbb{F}$, called the \emph{base field}, and the vectors can, for intuition building, be thought of as elements of $\mathbb{F}^n$ for some positive natural number $n$. It is possible to generalize this to things like $\mathbb{F}^\mathbb{F}$, the space of all functions from $\mathbb{F}$ to $\mathbb{F}$ but thinking about $\mathbb{F}^n$ is fine. To be even more concrete, you may imagine $\mathbb{R}^3$. But we are in the abstract algebra chapter here, so let's keep it general. Actually, even $\mathbb{F}^n$ is too concrete. We will instead just use $V$ to denote the set of vectors and $S$ to denote the set of scalars. 

% See also:
% https://claude.ai/chat/b8d78e4c-55b3-4baa-89b8-911f75022e2c
% https://claude.ai/share/db34436e-0da2-42d4-86ff-16ea73181767  public share link

% https://claude.ai/chat/b8d78e4c-55b3-4baa-89b8-911f75022e2c
% https://claude.ai/share/db34436e-0da2-42d4-86ff-16ea73181767
% https://claude.ai/share/db34436e-0da2-42d4-86ff-16ea73181767


\paragraph{Axioms} Like with groups, rings and fields, we will start with a bunch of axioms. We start with our set of vectors $V$ and the set of scalars $S$. We assume that we have a binary operation $+: V \times V \rightarrow V$ which we call vector addition and a binary operation $\cdot : S \times V \rightarrow V$ which we call scalar multiplication. The vector space axioms state that for any $\mathbf{u,v,w} \in V$ and any $\alpha, \beta \in S$, the following laws must hold true: ...TBC...

% ToDo: Give a list of vector space axioms, formatted the same way as the group axioms

% Linear algebra in hyperbolic space: Introduction to Gyrovector spaces
% https://www.youtube.com/watch?v=TMZmXT5IOsY at 0:20

\paragraph{Notation and Terminology}
As you have already seen, we will denote our set of scalars by $S$ our set of vectors by $V$ and we will typically use a lowercase boldface latin letter $\mathbf{v}$ to denote an element from $V$. When we need more vectors, we'll typically use letters around $\mathbf{v}$ like  $\mathbf{u}$ or $\mathbf{w}$. For the scalars, we will typically use lowercase greek letters like $\alpha, \beta$ and especially in the context of eigenvalues also $\lambda$. This is the same notation that you already know from the chapters about linear algebra earlier in the book and which is also widely used in the literature. ...TBC...TODO: List some other notations that other authors use.

% TODO: Make a section about notation and list all the different ways for denoting vectors and scalars that are in common use. I think, in linear algebra, we tend to use normal text for scalars and boldface for vectors. Some authors also use an arrow on top of vectors or an underbar. But are these notations also common in abstract alegbra? I think, there i see more use of: vectors: normal latin letters like v, scalars: greek letters like alpha. But maybe also things like $a v$ or $a X$

%...TBC... Explain that we use the boldface notation for vectors that the reader is already familiar with from the chapters about linear algebra. As an additional reminder that we are dealing with vectors, we'll often use the lette $v$ and the letters around it like $u,w$ to denote vectors.

% TODO: Maybe switch to the convention of using normal (non-bold) letters u,v,w for vectors and greek letters for scalars. When we later introduce algebras of quaternions and the like, the vector notation seems to be not really suitable anymore. But maybe we should mae that switch not here but in the section about algebras.

%===================================================================================================
\subsection{Subspaces} 
We have seen subgroups, subrings and subfields. It may not be surprising that we define a subspace of a given vector space in an entirely analogous way. A subset $U$ of a given vector space $V$ is called a \emph{subspace} of $V$ when it is itself a vector space. ...TBC...

% https://en.wikipedia.org/wiki/Linear_subspace

% https://math.stackexchange.com/questions/594887/notation-for-subspaces


\subsubsection{Types of Subspaces}

\paragraph{Lattices}
% A lattice is a set of discrete points within a vector space that is itself a vector space.
% BUT: there is also another notion of a lattice, namely an algebraic structure that follows laws similar to those of boolean algebra. It's a mess!

% Counting points on the E8 lattice with modular forms (theta functions) | #SoME2
% https://www.youtube.com/watch?v=AxPwhJTHxSg	


\paragraph{Hyperplanes}
% A hyperplane of dimension $k \leq n$ that passes through the origin of $\mathbb{R}^n$ is a lower dimensional subspace of the vector space $\mathbb{R}^n$.


%===================================================================================================
\subsubsection{Operations}

% We already know how to perform operations (like addition or scalar multiplication) on individual elements of a vector space, i.e. on individual vectors. There are also operations that operate on whole subspaces at once. These operations may produce subspaces as outputs or, more generally, sets with a certain anatomy.

%These operations result in sets




\paragraph{Orthogonal Complement}

% https://en.wikipedia.org/wiki/Orthogonal_complement


% Intersection
% Union - does not give a vector space, I think

\paragraph{Direct Sum}

%https://en.wikipedia.org/wiki/Direct_sum



\paragraph{Quotient Space}
Let $V$ be a vector space and $U$ be a subspace of $V$. We partition the vector space $V$ into equivalence classes by saying that two vectors belong to the same equivalence class, when their difference is an element of a certain subspace $U$ of our original vectors space $V$. ...TBC...

% https://www.youtube.com/watch?v=d6ixmHDlrNo&list=PLdTL21qNWp2Z2iZOktAgucHyWMnJPA6eu&index=23&t=4s

% Other operations:
% -Sum (different from "direct sum"),  S + T = { s + t | s in S, t in T }
% -Union?
% -Intersection, S \cap T = { u | u \in S \wedge u \in T }
% -Set product? S \times T = { (s,t) | s in S, t in T }
% -Disjoint union? Could this be similar to the (direct) sum?
% -Complement: \overline{S} = { v \in V | v \notin S } = { v | v \in V \wegde v \notin S }



\subsubsection{Implementation}
The operations on or between vector spaces, like taking an orthogonal complement or forming a direct sum, are very abstract ideas indeed. To get a more hands-on feeling for what is going on, I'll suggest a possible concrete (maybe naive) implementation of these operations. 

\paragraph{The Data Structure}
Firstly, we need a way to represent a vector space as a data structure. Every vector space has a basis, so it seems to make sense to represent a vector space by its basis. A basis is just a set of vectors. We will assume here that we represent a basis of a vector space as an $m \times n$ matrix $\mathbf{B}$ whose columns are considered to be our basis vectors. The matrix has $n$ columns, so we represent an $n$-dimensional space. The number $m$ - the number of rows of the matrix - can be any number $m \geq n$ but for the operations to be nontrivial, we'll want $m > n$. We may imagine $m$ to be the dimensionality of some embedding space. That is: we want to view our vector space of interest as a subspace of some possibly higher dimensional embedding space. This is because we want to implement operations like the orthogonal complement which is defined on a subspace with respect to a higher dimensional embedding space. It's a bit like with the set complement. It only makes sense with respect to an (perhaps implicitly understood) universal set.

%over the field $\mathbb{R}^n$ by an $n \times m$ matrix.

\paragraph{The Algorithms}
Given a vector space represented by its basis vectors stored in the rows of an $m \times n$ matrix $\mathbf{B}$, we now want compute a matrix $\mathbf{C}$ that represents a basis for the orthogonal complement of $\mathbf{B}$ with respect to the embedding space. That is, we want to compute a basis for $\mathbf{C} = \mathbf{B}^\perp$. The shape of $\mathbf{C}$ has to be $m \times k$ where $k = m - n$ because $\mathbf{B}$ represents a basis for an $n$-dimensional subspace of the $m$-dimensional embedding space, so the number of remaining dimensions is $m-n$. ...TBC...

% Give an algorithm that checks if some subspace V is a subspace of another subspace W. For example
% the embedding space may be U = R^3 (U for "universal space" or "universe"), W may be a plane 
% through the origina and V be a line thorugh the origin. It may or may not lie in the plane. The
% algorithm should figure that out.

%---------------------------------------------------------------------------------------------------
\subsubsection{Homomorphisms}
Just like groups and rings, vector spaces do also have a notion of homomorphisms. Such \emph{vector space homomorphisms} are also known as \emph{linear transformations} or \emph{linear maps}. If we view the transformation as a function $f: V \rightarrow W$ from some vector space $V$ so some (possibly) other vector space $W$, then the function $f$ must satisfy:
\begin{equation}
  f(\mathbf{v_1} + \mathbf{v_2}) = f(\mathbf{v_1}) + f(\mathbf{v_2}), \qquad
  f(\alpha \mathbf{v_1}) = \alpha f(\mathbf{v_1})
\end{equation}
for all $\mathbf{v_1, v_2} \in V$ and all $\alpha \in \mathbb{F}$, i.e. the $\mathbf{v_1, v_2}$ are vectors from $V$ and $\alpha$ is any scalar from our field of scalars. These two conditions are called \emph{additivity} and \emph{homogeneity} and taken together, they constitute the important notion of \emph{linearity}. Note again that on the left hand side, we are dealing with the addition and scalar multiplication of $V$ whereas on the right hand side, we are dealing with the operations of $W$. A simple example would be: $f: \mathbb{R}^2 \rightarrow \mathbb{R}^3$ with $f((x,y)) = (2 x + y, 3 x - 2 y, x - y)$. A slightly more advanced example dealing with infinite-dimensional vector spaces would be the derivative of a function. We could denote this as an operator $T: C^1(\mathbb{R}) \rightarrow C^0(\mathbb{R}) $ with $T[f] = f'$.

\medskip
Further important features of linear maps that follow from this definition are: (1) The zero vector of $V$ is mapped to the zero vector of $W$: $f(\mathbf{0}_V) = \mathbf{0}_W$. (2) The image of $V$ under $f$ is a subspace of $W$. (3) The kernel of $f$, i.e. the set of all vectors $\mathbf{v} \in V$ that get mapped to the zero vector, is a subspace of $V$. (4) When $f$ is injective such that we can invert it, then the inverse map is also linear.

%Furthermore $f[V] \subseteq W$

% Lineare Abbildungen und die Kern-Bild-Formel
% https://www.youtube.com/watch?v=blroCEbv1X8  by Weitz

% linear maps mus map the zero vector of V to the zero vector of W

% Kern-Bild-Formel aka Rangsatz aka Dimensionssatz
% dim V = dim ran(f) + dim ker (f)
% https://en.wikipedia.org/wiki/Range_of_a_function
% https://en.wikipedia.org/wiki/Kernel_(algebra)



%\subsection{Vector Spaces with Additional Structure} 
\subsection{Special Vector Spaces} 
Some vector spaces have additional structure built in by being equipped with additional operations. Here we will consider some operations that take one or two vectors as input and produce a scalar as output. If these operations satisfy certain requirements, we will have special vector spaces that are important enough to have their own names ...TBC...

\subsubsection{Metric Vector Spaces}
In a metric space, we have a function that takes two that takes two vectors $\mathbf{u,v}$ as inputs and produces a non-negative scalar as output. The geometric interpretation of the output of that function is the distance $d$ between the two input vectors which is why it is usually denoted by $d = \dist(\mathbf{u,v})$. A function that measures distances in a vector space is also called a metric. ...TBC...

%In a metric space, we have a distance function, also called a metric, that takes in two vectors and produces a scalar. ...TBC...

% The geometric interpretation of that function

\subsubsection{Normed Vector Spaces}
In normed vector spaces, each vector has a so called norm. A norm is a function that takes one vector as input and returns a scalar. ...TBC...

% https://en.wikipedia.org/wiki/Normed_vector_space
% -Has nice Venn diagram. Maybe reproduce it. Maybe replace the "Topological space" by "Vector 
%  space". Maybe the topological space is yet more general that the vector space? Figure out! Maybe
%  introduce another inclusion. But maybe don't draw a Venn diagram but instead just use a set
%  inclusion chain:
%  inner product spaces < normed spaces < metric spaces < vector spaces < topological spaces.
%  where the < is replaced by \subset


% I think, a metric space is more general than a normed space because each metric induces a norm by
% norm(x) = dist(x,0). Verify! Nope! It's the other way around! A norm induces a distance via: 
% dist(x,y) = norm(y-x). The difference y-x is always defined in a vector space. So yeah, in fact
% metric spaces are more general because for every normed space, we can immediately produce a 
% distance function via that formula. But we should discuss the idea of defining 
% norm(x) = dist(x,0). Explain why it doesn't work this way.

 

% https://en.wikipedia.org/wiki/Normed_vector_space


\subsubsection{Inner Product Spaces}
In an inner product space, we have the notion of an inner product which is a function that takes two vectors and produces a scalar. In this regard, an inner product is similar to a metric but the geometric interpretation of the output is different. An inner product measures the similarity of the directions in which the two vectors point, i.e. the angle between the two vectors. But for that interpretation to work, the output needs to be renormalized (i.e. divided by) the norms of the inputs. So, does that mean that we require our vector space to be a normed space in the first place? Well, yes and no. The way this works is that the inner product itself induces the norm as follows: In an inner product space, the norm is defined via an inner product of a vector with itself. ...TBC...verify!

% ToDo: explain orthogonality as a sort of independence of directions.

% Verify: Maybe we need to take the inner product of v its complex conjugate in the case of complex valued vectors. But Wikipedia seems to say otherwise. Check other sources.

% Does it make sense to use a different norm within an inner product space, i.e. one that isn't induced by the norm?

% https://en.wikipedia.org/wiki/Inner_product_space
% https://en.wikipedia.org/wiki/Inner_product_space#Basic_results,_terminology,_and_definitions


%\subsubsection{Banach Spaces}

% Banach spaces are complete, I think, 

%\subsubsection{Sobolev Spaces}

% Sort them from more general to more special

% What about topological spaces? But i think, they are not necessarily vector spaces

%\subsection{Generalizations} 
\subsection{Modules as Generalization}
If we replace the base field $\mathbb{F}$ with a more general ring $R$ in our definition of a vector space, we arrive at the definition of a module. Modules can therefore be seen as a generalization of the idea of a vector space that have less structure built in. In particular, we do not have multiplicative inverses in the set of scalars anymore and we also do not assume commutativity of the scalar multiplication anymore. ...TBC...


% Maybe also make a section about "almost vector spaces", i.e. structures with less structure
% Maybe call it generalizations
% https://en.wikipedia.org/wiki/Module_(mathematics)



\begin{comment}

-discuss operations like orthogonal complement, direct sum of vector spaces etc.
-maybe let Orthogonal complement and Direct sum be subsections, i.e. scrap the level "Operations"
-Implementation could also be a subsection
-Moduln/modules: like a vector space but made from an underlying ring rather than a field
-Make a section about additional structure due to norms, metrics, inner products, etc.



Was sind Vektorräume?
https://www.youtube.com/watch?v=DhW-Qef7mhQ
-at around 20:00 - some rules that follow from the axioms





\end{comment}