\section{Algebras}
In the previous section, we have seen that we can give additional structure to vector spaces by introducing a couple of more operations. The metric and the inner product were operations that took two vectors and produced a scalar, the norm was an operation that took one vector and produced a scalar. In this section, we will be concerned with operations that take two vectors $u,v$ and produce another vector $w$. This new operation is the one that we want to interpret as a \emph{product} between the two vectors. We'll denote the product between two vectors $u,v$ generically by $u \cdot w$ or sometimes shorter by $u v$. In certain special contexts, other symbols may be used such as $u \times v$. In abstract algebra, a vector space that is equipped with such a product is called an \emph{algebra}. In the context of discussing algebras, I will not use the boldface notation for vectors anymore for reasons that will be explained soon. ...TBC...





% Point out that we should not confuse it with the dot product. Maybe do not use the dot.

% Maybe point out the parallel between "an algebra" and "(abstract) algebra" as a field and "a topology" and "topology" as a field.

%  $\mathbf{u \cdot v}$

\paragraph{Axioms} 
Our minimal requirements for such a product operation are encapsulated in the following axioms:

\medskip
\begin{tabular}{l l}
Left distributivity:  & $u \cdot (v+w) = u \cdot v + u \cdot w$     \\	
Right distributivity: & $(u+v) \cdot w = u \cdot w + v \cdot w$     \\
Left linearity:       & $(\alpha u) \cdot v = \alpha (u \cdot v)$   \\
Right linearity:      & $u \cdot (\alpha v) = \alpha (u \cdot v)$
\end{tabular}
\medskip

The two linearity requirements can be (and often are) stated as the single requirement of \emph{bilinearity} which generally means that a binary operation must be linear in both of its arguments. That means that scaling one of the inputs results in an equal scaling of the output. Just like in rings, we need two distributive laws because we do not in general require our product to be commutative. It's also noteworthy\footnote{It is so noteworthy because in other contexts, associativity is usually the most basic one of our requirements.} that there is no general requirement of associativity and we will indeed encounter some non-associative algebras later on.

...TBC...

% TODO: explain that the two linearity requirements are often stated as one which is then called bilinearity. Maybe mention a few other ways to condense the axioms to less equations.


% Mention that in the last two lines we are dealing with two kinds of multiplication: the scalar multiplications and the vector multiplication.

%If this new binary operation on vectors happens to satisfy our usual distributive laws of multiplication over addition (recall that vector addition is already well defined), then we can interpret this new operation as a multiplication or product. 

% TODO: Explain requirements on the product (bilinear, left- and right-distributive, etc.). I think, bilinarity means (\alpha \vec{v}) \vec{w}

% TODO: Compare the definition of the term algebra with the definition of a ring. Does it make sense to say that algebras are special rings where the underlying set happens to be a vector space rather than just any set? What role do the scalars play here? Couldn't we just somehow embedd them in the set of vectors? In the case of matrices, the scalars can be represented by diagonal matrices with the scalar on the diagonal. In the case of complex numbers, we just leave the imaginary part at zero. Is such an embedding always possible? Try to find counterexamples. Maybe R^3 with the cross-product? Maybe graded algebras like tensor algebra or exterior algebra?
% -> rings are always associative
% -> in rings, the underlying set is not necessarily a vector space. It can be any group (the inherited group operation becomes the ring addition)

% https://en.wikipedia.org/wiki/Algebra_over_a_field


% These special vector spaces are important enough


\paragraph{Notation and Terminology}

% TODO: Explain why the boldface notations for vectors has beed dropped. 

% Vectors: u,v,w or x,y,z, sometimes v,w (complex) or q,p (quaternions)
% Scalars: \alpha, \beta, \gamma
% Vector components: a,b,c,d
% Multiplicatve indentity vector / identity element: e or 1 or I
% Zero vector: 0 ...but how do we distinguish it from the zero scalar?


% The reasons are:
% - This is common in the literature
% -
% - In many specific algebras, the property of "being and element of a vector space" tends to get
%   de-emphasized because the objects start to behave more like numbers again even though they are
%   still technically vectors (verify). Especially in the complex numbers, we tend to use z,w, in 
%   the quaternions we tend to use q,p, with matrices we tend to use uppercase bold. So, the 
%   lowercase bold notation is not so universal anymore in many specific algebras so doesn't seem to 
%   make too much sense to insist on it in a more general abstract discussion.
% - In many algebras the distinction between scalars and vectors gets a bit blurry anyway because 
%   more often than not, the scalars are embedded into the vectors space in some way (see section
%   about unitary algebras)
% - We want to reclaim the \cdot for denoting the multiplication. In conjunction with the typical
%   lowercase bold notation for vectors, readers would be very likely to mistake this for the
%   dot product (aka scalar product) where we actually mean the algebra specific product that takes
%   two vectors and produces a third vector.



%===================================================================================================
% Old:

%\subsection{Additional Structure from Products} 

%There are vector spaces that have additional operations defined on the vectors. These spaces are classified based on the properties of these additional operations. Typically, there is at least one additional binary operation between vectors, i.e. an operation that takes two vectors as input and produces another vector as output. This operation is often interpreted as "product" between the vectors although it may be any operation - remember that we are doing abstract algebra here. It could be a commutator between matrices (that happens in Lie algebras), a concatenation of loops (that happens in algebraic topology), etc. The product operation is typically bilinear. Vector spaces with such a bilinear product are called \emph{algebras over a field} or just \emph{algebras}\footnote{The term "algebra" is really overloaded a lot in mathematics. If you find that confusing then welcome to the club!}.

% algebra as a field, an algebra as a vector space, a sigma-algebra, algebra as in signatures and algebras, ...

% These vector spaces with additional structure are called \emph{algebras over a field} or just \emph{algebras}\footnote{The term "algebra" is really overloaded a lot in mathematics. If you find that confusing then welcome to the club!} when the additional structure arises from a bilinear product operation.

% Maybe make the subsection "Extra Structure" with subsubs
% Products, Gradation, Derivation, Norm, Inner product

%an additional operation defined that serves the purpose of a product between vectors that yields another vector. Depending on the properties of that product operation, they get different names.

% https://en.wikipedia.org/wiki/Algebra_over_a_field


\subsection{Motivating Examples}
%\subsection{Some Algebras over $\mathbb{R}$ as Motivating Examples}
%\subsection{Some Algebras over $\mathbb{R}$}
Abstraction is best understood when we first see a bunch of concrete examples to motivate ourselves to have a bunch of different structures that have enough in common to make certain abstractions worthwhile. So will now look at a couple of motivating examples of algebras over the real numbers $\mathbb{R}$, which we also call $\mathbb{R}$-algebras.


\subsubsection{The Complex Numbers}
We are already used to the idea to think of complex numbers as points in the real $2D$ plane, i.e. to envision $\mathbb{C}$ as the vector space $\mathbb{R}^2$ with some additional structure. This "additional structure" is brought to us via definition of the product between two complex numbers. This is the new operation that was totally unheard of in the more frugal world of plain $\mathbb{R}^2$. We cannot (yet) multiply $2D$ vectors, let alone divide them or even do more crazy stuff like raising one to the power of another. All of these things and more will become possible thanks to a simple definition. Let's start from identifying two vectors $(a,b)$ and $(c,d)$ $\in \mathbb{R}^2$ with the complex numbers $a + \i b$  and $c + \i d$ respectively where $\i$ is defined by the property that $\i^2 = -1$. The goal is to find expressions for the two vector components of their product. Of course, we already know the result. They must be the real and imaginary part of the complex product. But for purposes of demonstrating the general approach, let's churn through the derivation step by step. When we have these expressions in hand, we have a product on $\mathbb{R}^2$. 

\medskip
\begin{tabular}{l l}
$(a,b) (c,d)$                        & Product written with vectors. Rewrite as complex product. \\
$(a + \i b) (c + \i d)$              & Use right distributivity\\
$a (c + \i d) + \i b (c + \i d)$     & Use left distributivity \\
$a c + a \i d + \i b c + \i b \i d$  & Use commutativity of multiplication  \\
$a c + \i a d + \i b c + \i \i b d$  & Use $\i \i = \i^2 = -1$  \\
$a c + \i a d + \i b c - b d$        & Use commutativity of addition  \\
$a c - b d   + \i a d + \i b c$      & Use left distributivity for factoring out $\i$  \\
$a c - b d   + \i (a d + b c)$       & Extract terms with and without factor $\i$  \\
$(a c - b d, a d + b c)$             & Result written as vector  \\
\end{tabular}
\medskip

I was very detailed and careful here while making at every step explicit, which law was invoked. If you are concerned about the fact that the "Use commutativity of addition" step also involves a subtraction, recall that the notation the $- bd$ is just a shorthand for $+ (-(bd))$. If you are concerned about the validity of the "Use commutativity of multiplication" step, then you have a point. The step is valid in the special case here, but in general we need to be careful with such things because such a commutative law of multiplication is not always given. We will come back to this when deriving the formulas for the quaternion product later. ...TBC...

% ToDo: write down the end result in various forms including row and column vector notation.



\subsubsection{The Hyperbolic Numbers}
The complex product is not the only possible way to define a product between two vectors of $\mathbb{R}^2$. Another possibility is ...TBC...ToDo: introduce $j$ with $j^2 = 1, j \neq 1$, define hyperbolic numbers as $a + b j$, multiply out to obtain the formulas for the 2 parts of the number

\subsubsection{The Dual Numbers}
There is yet a third way to define a product between two vectors of $\mathbb{R}^2$. ...TBC...ToDo: introduce $\varepsilon$ with $\varepsilon^2 = 0, \varepsilon \neq 0$, define dual numbers as $a + b \varepsilon$, multiply out to obtain the formulas for the 2 parts of the number



%\subsubsection{Vectors in $3D$ and the Cross Product}
%\subsubsection{Vector Algebra in $3D$}
\subsubsection{Vectors in $\mathbb{R}^3$}
ToDo: Introduce the cross product, explain its properties (anticommutative, non-associative, ...), explain why this makes this algebra a bit awkward to work with, hint at the wedge product as solution, explain that a cross-product with certain desired properties (list them) can exist only in 3 and 7 dimensions

% Works only in 3D, Maybe we can call it 3D vector algebra

% why there is no four dimensional cross product.
% https://www.youtube.com/watch?v=FPoo7rm0OGI
% -Has some expectations that we have on a cross product for what properties it should have


%\subsubsection{Vectors in $\mathbb{R}^n$}
% With vectors in R^n, we can build an algebra using the (naive) component-wise product. It is indeed associative, distributive over addition and even commutative. The neutral element is a vector of all ones and the inverse elements are the evctors of the component-wise reciprocals. It's not a division algebra though because even one zero element in a vector makes it non-invertible and the unique neutral element of addition, i.e. the bona fide zero element of the algebra, is the vector of all zeros. Vectors containing zeros but also non-zero elements are the zero-divisors, I guess. Practical examplesof use: in programming (for example, in NumPy arrays, I think. In MatLab/Octave, it'S the .* operator)


\subsubsection{Matrices in $\mathbb{R}^{n \times n}$}
ToDo: Remind the reader that $n \times n$ matrices over $\mathbb{R}$ already constitute a bona fide vector space and observe that we also already have matrix multiplication defined and when both matrices are of the same quadratic $n \times n$ shape, then so will be their product, such that the space is closed under this product. This product satisfies the other requirements as well (bilinearity, etc.), so we have an algebra. ....

% TODO: List a couple of relevant sub-algebras: rotations (orthogonal (or unitary) matrices), symmetric (or Hermitian) matrices -> I think, the product is then commutative, maybe reflections or permutations could also be interesting? Maybe triangular matrices, diagonal matrices (would resmble vectors with a component-wise product).

% TODO: Maybe also build algebras with other matrix products like component-wise, Hadamard, ...

% https://en.wikipedia.org/wiki/Hadamard_product_(matrices)
% https://de.wikipedia.org/wiki/Hadamard-Produkt

\subsubsection{Tensors in $\mathbb{R}^n$}
ToDo: Introduce tensors as generalization of scalars, vectors and matrices, introduce the tensor product, explain how the tensor product produces objects of different rank such that the tensor algebra is a graded algebra (verify), ...

% Are grade and rank synonymous?

\subsubsection{The Quaternions}
ToDo: Introduce quaternions as a 4D division algebra over $\mathbb{R}$. Define the quaternion product, give formulas for inversion, norm, etc. list important identities, give a bit of the historic backstory of how Hamilton discovered them while his goal was to find a 3D division algebra, explain various representations and their pros and cons, 


% Another Imaginary Unit
% https://www.youtube.com/watch?v=6w0Qb7rLUa0


\subsubsection{The Wedge Product}
% Similar to cross product but generalizes to nD


% ToDo: dual numbers, hyperbolic numbers, octonions, sedenions


\subsection{Classification of Algebras}
Okay - we have now seen a big enough zoo of different algebras to be ready for some abstractions. We can classify different types of algebras by abstractly requiring the product to satisfy certain additional properties and give these types special names.

% TODO: explain how in certain algebras scalars can be embedded into the set of vectors such that we do not necessarily need them as separate data type anymore ... for matrices, we just form the diagonal matrix with the scalar in the diagonal, for complex numbers, we just keep the imaginary part zero. But I think, this is not possible with all algebras. For example, in tensor algebra, we still need them. Maybe also in 3D vector algebra. I'm not sure if there is already a name for it. If not maybe mention it as another possible classification - "scalar embeddable algebras" or something


\subsubsection{Commutative Algebras}
Very unsurprisingly, an algebra is called commutative when for all elements $u,v$ of the underlying set of vectors the product $uv$ is equal to the product $vu$. ...TBC...

% TODO: Introduce the commutator  [a,b] = ab - ba  as a measure for the (non) commutativity of a,b

% TODO: Give examples of commutative algebras: numbers (N,Z,Q,R,C), polynomials, functions, symmetric and Hermite matrices (verify!), mention that the "ideal" is a central concept in the subfield of "commutative algebra" - yeah - first mention that this is actually a whole subfield formerly known as "ideal theory"

% Maybe make a section about non-commutative algebras and give some examples for some other rules (like anti-commutativity) that could hold instead. Maybe this is also the right context to introduce the commutator


\subsubsection{Unitary Algebras}
A unitary algebra is an algebra where a multiplicative identity element $e$ exists. This element $e$ satisfies:
\begin{equation}
 e \cdot v = v, \quad v \cdot e = v \qquad \forall v \in V
\end{equation}
A noteworthy feature of unitary algebras is that the set of scalars (which is inherited from the underlying vector space) can be embedded into the set of vectors by mapping any scalar $\alpha$ to the vector $\alpha e$ where $e$ is the multiplicative identity element. Thus, in unitary algebras, the scalars become superfluous because their effect on vectors (of scaling them) can also be achieved by certain special elements of the vector space itself via the additional multiplication operation that the algebra provides. The additional scalar multiplication that is inherited from the vector space becomes redundant. For example, in the algebra of complex numbers, the scalars are embedded as those complex numbers whose imaginary part is zero. The real numbers are embedded in the complex numbers.  In the algebra of $n \times n$ matrices, the scalars are embedded as the diagonal matrices where all diagonal elements are equal to the given scalar. But there are also algebras where the scalars cannot be embedded into the set of vectors. Take $\mathbb{R}^3$ with the cross product as counterexample. This algebra is non-unitary because no vector $e$ exists such that the cross product $e \times v$ with it leaves an arbitrary other vector $v$ as is. ...ToDo: Explain that this is the simplest nontrivial Lie algebra (Claude AI said this).


\subsubsection{Division Algebras}
A division algebra is an algebra in which we can always divide by an element $a$ unless $a = 0$. ...TBC...

%A division algebra is an algebra in which each element $a$ has a multiplicative inverse. ...TBC...
% Nope! That may be wrong! We may have to distinguish between left and right division whne the algebra is non-commutative. I think multiplicative inverses need onyl exist in commutative division algebras (verify!)

% what about zero divisors?

% https://en.wikipedia.org/wiki/Division_algebra
% -Associative division algebras have no nonzero zero divisors


\subsubsection{Associative Algebras}
A associative algebra is an algebra in which the associative law holds for the multiplication. ...TBC...
% Maybe mention that the associative rules is supposed to hold for addition anyway.








\subsubsection{Non-Associative Algebras}
% https://en.wikipedia.org/wiki/Non-associative_algebra
Certain types of non associative algebras satisfy weaker forms of associativity and some of these are noteworthy enough to have names. We'll look at some of them

% Maybe there could also be alternative rules to associativity that are not necessarily weaker forms of it...maybe something like anti-associativity analoguous to anti-commutativity? Is that a thing? Could we define an "associator" analoguous to the commutator? Like [a,b,c] = (ab)c - a(bc)

\paragraph{Alternative Algebras} 
%https://en.wikipedia.org/wiki/Alternative_algebra

\paragraph{Flexible Algebras} 
% https://en.wikipedia.org/wiki/Flexible_algebra

\paragraph{Power Associative Algebras}
% https://en.wikipedia.org/wiki/Power_associativity

% ToDo: explain power-associativity as a weaker form of associativity. (see "Zahlen" (Ebbinghaus), 
% pg. 151). x^m x^n = x^(m+n). In general, we cannot assume x x^2 = x^2 x even in the commutative
% case, so power associativity demands that to be satisfied


% https://en.wikipedia.org/wiki/Jordan_algebra


\subsubsection{Lie Algebras} 
% https://en.wikipedia.org/wiki/Lie_algebra
% https://en.wikipedia.org/wiki/Lie_bracket_of_vector_fields


% https://en.wikipedia.org/wiki/Lie_group%E2%80%93Lie_algebra_correspondence

%\subsubsection{Vertex Algebras} 


\subsubsection{Poisson Algebras} 
% https://en.wikipedia.org/wiki/Poisson_algebra
% https://en.wikipedia.org/wiki/Product_rule
% https://en.wikipedia.org/wiki/Derivation_(differential_algebra)


\subsubsection{Differential Algebras} 
Differential algebras have additional univariate operator that we call a \emph{derivation}. That is: a derivation takes one vector as input and produces another one. Let's denote the operation as $D$ and use functional notation, i.e. $D(a)$ means: apply the operator $D$ to the vector $a$. The operator must satisfy the product rule:
\begin{equation}
D(a \cdot b) = a \cdot D(b) + D(a) \cdot b
%D(a \times b) = a \times D(b) + D(a) \times b
\end{equation}
Compare this to the product rule that you know from calculus and you will notice that it is indeed the same rule. The vector space of differentiable functions with the pointwise multiplication between the functions as product and the derivative as derivation does indeed form a differential algebra. A subspace of that is the space of all polynomials. If we represent the polynomials as sequences of coefficients, we could use convolution as our product operation. The derivation operation would be to apply the rule how to differentiate polynomials: left-shifting the sequence $a_n$ by one and multiplying by the old index, i.e. informally $D(a_n) = n a_{n+1}$. If we now forget that the sequences originally were meant to represent polynomial coefficients, we have now a differential algebra on (finite) sequences of numbers without necessarily having to think about functions. It would straightforwardly generalize to infinite sequences, too - even if they don't converge as a power series, i.e. don't define a function [VERIFY!]. ...TBC...what other interesting examples are there (subalgebras don't count)? 

% TODO: Maybe use $f,g$ instead of $a,b$ because that is more suggestive. But mention why we do this here.

% TODO: Emphasize the order of D and mul on the right hand side. It matters in non-commutative algebras. Figure out and explain why the definition uses this order and not any of the other 3 possibilities. The 4 possibilities are: a D(b) + D(a) b, a D(b) + b D(a), D(b) a + D(a) b, a D(b) + b D(a), 

% Maybe use \times instead of \cdot to denote a general product in this section and maybe treat it as optional, i.e. a b means the same as a \times b

% Sequences of numbers with convolution as multiplication. the derivation is: left-shifting by one and dividing by the old index, i.e. b_k = (a_{k+1})/(k+1). That's how polynomial multiplication and taking derivatives work. But we do not need to interpret the sequence as sequnece of polynomial coeffs

% My C++ code says that the commutator in geometric algebra obeys the product rule - see enum class ProductType in Tools.cpp, line 5661

% There is a univariate operation that satisfies the product rule
% https://en.wikipedia.org/wiki/Derivation_(differential_algebra)

% What about the arithmetic derivative?

% I think  computer algebra systems rely on such an "algebraization" of calculus to do their computations

% https://en.wikipedia.org/wiki/Gerstenhaber_algebra
% https://en.wikipedia.org/wiki/Superalgebra



\subsubsection{Graded Algebras} [VERIFY!]
In a graded algebra, we are dealing with objects of different kinds that cannot be mixed up with one another in addition but can be in multiplication. As an example, consider the algebra of tensors over a vector space $V^n$. The product is, of course, the tensor product. We cannot add tensors of different ranks $m$ and $n$ but we can always multiply such tensors to get a new tensor of rank $m+n$. The rank of the tensor may also be called the \emph{grade}. Another example is the exterior algebra with the exterior product. We have scalars, vectors, bivectors, trivectors, etc. We cannot\footnote{Not \emph{yet} in exterior algebra (aka Grassmann algebra), that is. But that problem is satisfyingly resolved in geometric algebra (aka Clifford algebra).} add a vector to a bivector - but we can multiply them to get a trivector ...TBC...

% Q: Is geoemtric algebra a graded algebra? We do have objects of different grades but wa can mix them additively.

%For example, it doesn't make sense to add a scalar to a vector but multiplication between a scalar and a vector does. 

%In a graded algebr

% Formally, the space we are dealignwith is expressed a a direct sum of the vector spaces for each grade

% https://en.wikipedia.org/wiki/Graded_ring#Graded_algebra

% https://en.wikipedia.org/wiki/Graded_ring
% https://en.wikipedia.org/wiki/Graded_vector_space

% https://de.wikipedia.org/wiki/Gra%C3%9Fmann-Algebra

% https://en.wikipedia.org/wiki/Graded_structure


\subsection{Existence of Algebras}
Having all of these possible classifications in place, we could now proceed to ask: "Which types of algebras are possible?". Could there be a $3D$ division algebra, for example? And that specific question is far from being contrived. A $3D$ division algebra was actually precisely the thing that Hamilton so desperately wanted to find. Instead, what he found were the quaternions which are $4D$. And there's a deeper reason for that ...TBC...

% We now know that a $3D$ divison algebra is provably impossible (verify - maybe it's only impossible when we also require some other constraints?) ...TBC...

% ToDo: Cite results for which types of algebras are possible and which are not. For example, there are only 3 division algebras over $\mathbb{R}$, namely $\mathbb{R}$, $\mathbb{C}$, $\mathbb{H}$. Explain how this knowledge could save one a lot of futile work (use Hamilton's search for a $3D$ division algebra as example).



% Why are there no 3 dimensional "complex numbers"?
% https://www.youtube.com/watch?v=L-3AbJM-o0g



%\subsection{More Examples}
\subsection{The Zoo of $\mathbb{R}$-Algebras}
We have already seen a bunch of the more common $\mathbb{R}$-algebras in the section about the motivating examples. Now that we have some terminology for the classification at our disposal, we are ready to look at the broader zoo of all the various different $\mathbb{R}$-algebras and sort them into their appropriate combination of categories. ...TBC... ToDo: Classify the algebras that were already seen, introduce more algebras along with their classification, maybe widen the perspective to algebras over other fields or rings.

% Maybe for an algebra ove a ring, we cannot use a vector space but instead need a module?


\begin{comment}

-Maybe begin a motivating sections using the 3D cross product and the 2D complex product and the 4D
 quaternion product.

-Maybe list a couple of hypercomplex algebras (biquaternions, ...). See "Mathematik mit 2x2 
 Matrizen". It gives a lot of examples.



-Matrix Lie groups are (?):
 -Vector space under matrix addition (just like more general sets of matrices)
 -Group under matrix multiplication (that's the special property?)
 
-NxN-matrices together with scalars form a vector space under matrix-addition (as the
 vector-addition operation). Can we use other matrix operations to form such a vector
 space? Multiplication? Commutator? -> figure out!
 I think, for the distributive law to hold: a * (B * C) = a*B * a*C for scalar a and matrices B,C, 
 we must have a^2 = a so the set of scalars can contain only elements that square to themselves? In
 Z_6, we have 0,1,3 as such elements - but Z_6 is not a field...hmm...but perhaps Z_2 could work.
	
A gentle description of a vertex algebra.	
https://www.youtube.com/watch?v=7j4YVIFmAXw	
-Associative algebra: 
 -A vector space with a (bilinear, associative) product
 -Examples: 
  -NxN matrices (with matrix multiplication), 
  -polynomials (with polynomial multiplication)
  -multivariate polynomials with infinitely many variables
 -It's kinda like a ring with some extra structure.
-Lie algebra:
 -The product is usually written as a "Lie bracket": [x,y]
 -The product not associative but it follows the Jacobi identity 
 -The product is anticommutative aka antisymmetric (verify!)
 -Examples:
  -Take any associative algebra and define [x,y] = xy - yx, i.e. use the commutator
...ToDo: continue watching at 9:55 - there's more stuff

alternative algebra -- featuring the octonions!
https://www.youtube.com/watch?v=ZC7YofZp-cw
-An alternative algebra also starts with a vector space with a product
-The product satisfies (aa)b = a(ab)  and  (ab)b = a(bb). Note that these rules would follow
 immediately from associativity (ab)c = a(bc), if we would have associativity (which we don't). So, the "alternative" rule is a weaker rule than associativity.
-The "associator" is an operation defined as [a,b,c] = a(bc) - (ab)c. Thisn operation is an alternating multilinear form if we have an alternting(alternative?) algebra
...ToDo: continue watching at 5:00


The Multiplication Multiverse | Infinite Series
https://www.youtube.com/watch?v=H4I2C3Ts7_w
-Lie-Algebra: Product is anti-commutative and satisfies the Jacobi identity
-Other possible identities: Poisson identity, Super-commutativity
 ...Poisson Algebra, Gerstenhabe algebra
-Topology gives a way to "multiply" loops - the multiplication is just concatenation

Associahedra: The Shapes of Multiplication | Infinite Series
https://www.youtube.com/watch?v=N7wNWQ4aTLQ&list=PLa6IE8XPP_glwNKmFfl2tEL0b7E9D0WRr&index=19



-Division algebra: An algebra where every element has a (unique?) multiplicative inverse.

-Commutative algebra: multiplication is commutative. See:
 https://en.wikipedia.org/wiki/Commutative_algebra
 Examples: 
 -integers, p-adic numbers, reals, ...
 -polynomials
 -Maybe symmetric matrices (when complex: Hermite?). We have (AB)^T = B^T A^T and when A,B are
  symmetric, T does nothing

-Lie algebra:

-Banach algebra: https://en.wikipedia.org/wiki/Banach_algebra

Some of my ideas to further classify algebras (todo: check, if these names are already in use):

-Exponentiation algebra. Defines another binary operation a^b that satisfies the usual power law:
 a^b * a^c = a^(b+c). Maybe some other laws should be satisfied as well but that seems to be the
 most important one that establishes a relation between all 3 operations (addition, multiplication,
 exponentiation).
 -Maybe try to define exponentation for matrices by first defining exponentiation 
  B = e^A = exp(A) via a power series, then defining the logarithm as inverse of that 
  A = log(B) iff B = exp(A) and then use the change of basis formula C = A^B = log(A) * exp(B).
  Exponentiation should satisfy e^A e^B = e^(A+B) ..but maybe only if e^A and e^B commute. For 
  non commuting matrices, there is a more general fomula involving commutators - the 
  Baker-Campbell-Hausdorff formula:
  https://en.wikipedia.org/wiki/Baker%E2%80%93Campbell%E2%80%93Hausdorff_formula
  In general: e^A e^B = e^C with 
  C = A + B  +  (1/2) * [A,B]  +  (1/12) * ([A,[A,B]] + [B,[B,A]]) + ...
  where [A,B] = AB - BA is the commutator between A and B
  Maybe we should also demand e^A e^-A = I and (e^A)^n = e^(nA) because that's what the Google 
  AI overview says when searching for: "general rule for exponential functions when multiplication 
  is not commutative"

-Composition algebra or functional algebra: Defines another binary operation which abstracts the
 idea of function composition. Maybe if the algebra is also already a differential algebra, it
 should also satisfy the chain rule for the operations of derivation and composition. Try it with
 polynomials and power series.

-Logical algebra: we have 2 binary operations and, or and one unary opration not and they satisfy
 the rules of logic. I think especially interesting are the mutual distributive laws. Maybe we 
 should call it "bidistributive". Aside from Boolean logic itself, set algebra with intersection,
 union and complement also follows these rules. It basically inherits them from the definitions of
 these set operations (which are in logical terms using and, or, not). Maybe a 3 valued logic also
 follows these? What about probability logic? What about Z or N with min(a,b), max(a,b) and maybe
 additionally with zero(a,b) = 0. I think, these are lattices. Why that name? Well, in abstract 
 algebra, the names are traditionally badly chosen. A "ring" has also not really anything to do
 with what we mean by the word in everyday language. We would expect it to be something that somehow
 cycles back on itself in some way - maybe like a cyclic group but the actual definition of a ring 
 has nothing to do whatsoever with such a cyclic behavior. For a lattice, I would assume something
 like the Gaussian integers. 

-What about the algebra of rational functions?

-What about natural numbers with gcd, lcm and perhaps also mul?


More ideas:

-Try defining a derivation operation on the space of NxN matrices. Maybe this operation can
 be expressed in the 5x5 case by multiplication with the fixed matrix:

        0 1 2 3 4 5              0 1 2 3 4 5
        0 0 1 2 3 4              0 0 2 3 4 5
   D =  0 0 0 1 2 3 ..or maybe   0 0 0 3 4 5  ?
        0 0 0 0 1 2              0 0 0 0 4 5
        0 0 0 0 0 1              0 0 0 0 0 5

  Motivation: Consider what such a matrix does to a vector (a0,a1,a2,a3,a4). What we want it to do
  is to: multiply by (old) index and then shift. This may turn the algebra of matrices in a 
  differential algebra, I think.
  See also:
  Doing calculus with a matrix!
  https://www.youtube.com/watch?v=2iK3Hw2o_uo


Questions:

- When we talk about an R-algebra, does the R there refer to the set of scalars or to the vector
  components? Do we even need to assume that the vectors hava a notion of components? There are
  vector spaces, where "components" are not really a thing - for example function spaces. Do they
  also count as algebras (over R)?

- How does linear algebra fit into this framework? Can we talk about "the" linear algebra over R?
  I think, generally, linear algebra as a math subfield deals with MxN matrices and the product
  between matrices is really only a partial function. It's only defined between MxP and PxN 
  matrices. We are not really dealing with a vector space here anymore because we cannot add
  matrices of different shapes - unless we assume M and N to be fixed once and for all. In this 
  case we do have a product, the matrix product, only when M=N. Although there may be other matrix
  products like the Kronecker product that may also work on heterogeneous matrices...but we still
  can't really add matrices of different shapes.

- How do algebras compare to rings? 1: rings are always associative, algebras not necessarily, ...

why some number systems can’t divide
https://www.youtube.com/watch?v=ekHoYfT3rFI
-About extension of the integers where we can still do division with remainder, so it's
 not about fields or division algebras
-Division with remainder is defined as follows: for all a,b, in Z there are q,r in Z such that:  
 a = b*q + r  with  0 <= |r| < |b|. Here q is called quotient and r remainder.
-Idea: Maybe in the integers, we can also work with negative remainders to make their absolute 
 values even smaller. Normally, we would say that 23/5 = 4r3 but how about using 5r-2 instead? In
 general, we could do floor division as usual to obtain q, compute the differences d1 = a - b*q
 and d2 = a - b*(q+1) and whatever difference has the smaller absolute value decides if we report
 q itself or q+1 as result (or quotient) with the corresponding remainder. I think, if b is even,
 we could get ties and we need a rule to break them. Maybe then choose the positive remainder and
 q itself rather than q+1 as quotient.
-Could we define such a division-with-remainder also for matrices? Like A = B*Q + R where A,B,Q,R 
 are all matrices? For regular (i.e. invertible) matrices A, the quotient Q would be the inverse of
 A and the remainder R would be zero but for singular matrices, we could perhaps find some
 pseudo-inverse that somehow minimizes some matrix-norm of the remainder matrix? Try to solve the
 minimization problem |A - BQ| = min where |...| is some matrix norm. Maybe try the Frobenius norm
 which is just the sum of the squares of the entries. Maybe also try just using the Moore-Penrose
 pseudo inverse for Q. Does this lead to something interesting and useful? Could it have something 
 to do with orthogonal projections? Maybe when A is singular, its columns are not spanning the 
 whole space R^n but only a lower dimensional subspace of it BQ is somehow the orthogonal distance
 between A and the subspace or something?
-Maybe we could also put additional constraints on what type of matrices are allowed, like for
 example, symmetric, pos. definite, orthogonal, integer, etc. and find a Q that minimizes the norm 
 of R subject to these constraints?
-Maybe we could also try to do this with an outer product rather than the normal matrix product.
 Like  A = u * v^T + R   where A is an MxN matrix and u,v are M,N vectors respectively. I think,
 we would try to express A as an outer product of two vectors, i.e. attampt a sort of tensor
 decomposition or rank decomposition (or whatever this is called). We would try to express a rank 
 2 tensor as tensor product of two rank 1 tensors ...plus the remainder because in general, this is
 not always exactly possible. If it's possible, it's actually exceptional because a full matrix has
 many more degrees of freedom than the 2 vectors, i.e. more dimensions. A 3x3 matrix has 9 entries, 
 whereas a pair of 2 3-vectors has only 6, for example




The unreasonable effectiveness of linear algebra.
https://www.youtube.com/watch?v=Rqv3cXt8ZNU
-Gives an example of a 4D function space with a derivative operator

Why can't you multiply vectors?
https://www.youtube.com/watch?v=htYh-Tq7ZBI


https://de.wikipedia.org/wiki/Vektorraum
https://de.wikipedia.org/wiki/Heterogene_Algebra
-includes vectros spaces

\end{comment}
