\section{Continuous Functions}

\subsection{Limits}
Consider the function $f(x) = \sin (x) / x$. As it stands, it is not defined at $x = 0$ because we would have a division by zero there. In fact, we have an indeterminate form of the form "zero divided by zero". We could just define the function to assume an arbitrary value at $x = 0$, say $f(0) = 0$. But it turns out that there is a more natural way to fill the gap. Imagine plugging ever smaller numbers into $f(x)$, like $0.1, 0.01, 0.001, $ etc. and observe what $f(x)$ does. It turns out that $f(x)$ will approach one. That may suggest to define $f(0) = 1$. But not so fast: What if we would approach zero from the left, using $-0.1, -0.01, -0.001,$ etc. as inputs instead? Good news: It turns out that $f(x)$ also approaches one in this case, so we are indeed justified to define $f(0)=1$ to fill the gap in a natural and meaningful way. The idea that a function approaches a certain value $f(x_0)$ when its input values approach a given value $x_0$ leads to the idea of a limit. We need to distinguish the limits approaching from the left and approaching from the right. Only when both of these are equal, we say that \emph{the} limit exists and we define it to be just that value. If the function $f$ has multiple inputs, we must actually consider all possible paths on which we could approach the point where $f$ is undefined. For the time being, we'll focus on the simpler 1D case where we can approach a point on the number line only from two directions.
TODO: introduce mathematical notation for limits and the epsilon-delta definition


% To define a limit mathematically, we consider a function $f(x)$

% https://en.wikipedia.org/wiki/Limit_(mathematics)
% https://en.wikipedia.org/wiki/Limit_of_a_function#(%CE%B5,_%CE%B4)-definition_of_limit


% https://www.youtube.com/watch?v=n2syaXR8y_k
% log-of-limit is limit-of-log - does that generalize to other functions?
% this is actually a good example:
% $\lim_{n \Rightarrow \infty} (\frac{\sqrt[n]{a} + \sqrt[n]{b}}{2})^n = \sqrt{a b} $

% Explain why lim_{x -> inf} f(x) / g(x) = 1  means the same as  lim_{x -> inf} (f(x)-g(x))/f(x) = 0
% ...these are the two ways of static asymptotic equivalence of f(x) and g(x) - I think

\subsection{Continuity}
If the limit of a function $f$ exists at a given $x_0$ and agrees with the actual function value $f(x_0)$, we say that the function $f$ is continuous at $x_0$. With that definition in place, we can now be more specific about what we have meant with "filling the gap in a natural way". We have filled it in such a way as to make the resulting function continuous at $x_0 = 0$. At all other values of $x_0$, it already has been continuous all along. If a function is continuous at \emph{all} values $x_0$, no matter what we choose as $x_0$, we say that the function is continuous without the qualification "at $x_0$". Intuitively, continuity means that the graph of the function has no sudden jumps (aka. discontinuities) and we can draw it without lifting the pencil or drawing vertical lines. TODO: give examples and counterexamples - include the Dirichlet function as example for a function that is nowhere continuous, include a piecewise definition, one with a pole like 1/x

% Laws for limits:
% linearity: lim (f(x) + g(x)) = lim(f(x)) + lim (g(x)), lim(a*f(x)) = a * lim (f(x))
% l'Hospital's rule



% Maybe include this:
% https://en.wikipedia.org/wiki/Limit_inferior_and_limit_superior#limit_superior
% but it applies to sequences, so maybe include it there.

% See also:
% https://en.wikipedia.org/wiki/Uniform_continuity
% https://en.wikipedia.org/wiki/Lipschitz_continuity
% https://en.wikipedia.org/wiki/Absolute_continuity






\begin{comment}

Can A Function Be Continuous At Only One Point?
https://www.youtube.com/watch?v=TdhUNhGbfD8

f(x) = 0 if x is irrational and x if x is rational. It's continuous only at x=0

f(x) = 0 if (x-1)*(x-2)*... if x is irrational and x if x is rational. continuous only at 1,2,...


The 7 Indeterminate Forms that Changed Math Forever
https://www.youtube.com/watch?v=X29K1EIT_Hw


https://www.youtube.com/watch?v=q_Cc4ebGCOE  Grenzwerte von Funktionen
- Discusses what it means for a function to have a limit of y0 at a given point x0. It means that 
  for *any* possible sequence of numbers that we cant construct that converges to x0, it must be the 
  case that the corresponding sequence of function values converges and converges to *the same* 
  value y0. He gives the counterexample f(x) = sin(1/x) which has no limit at x0 = 0. He constructs
  sequences with limit x0 = 0 for which the corresponding sequence of function values coverges to 0,
  to +1, to -1 and even one that oscillates between -1 and +1. But to introduce limits like this, we
  may need to put the section about sequences before this section. Look if that's doable in the 
  sense that the section does not refer to any material that will then come after it. If so, it may
  be a viable option to change the order in that way. ..yeah - it actually does make sense - except 
  for the convolution integral that appears there. But we may just put a star on it - it's just a 
  side note and not relevant for the rest of the section.


These Limits Are Too Complicated for Calculus
https://www.youtube.com/watch?v=-uIwboK4nwE


Why 1^inf Isn’t 1?
https://www.youtube.com/watch?v=gaMd7hP8K54



Do we even need the real numbers??
https://www.youtube.com/watch?v=vV7ZuouUSfs
- Shows that continuity is troublesome because a step function with a step at an irrational number
  would actually qualify as continuous over the rationals. The step discontinuity is "no seen" in a
  way.
- Shows that on a calculus that only has rational numbers to work with (that means the domain and 
  codomain are the rationals and we only look at rational functions), the intermediate value 
  theorem doesn't hold anymore.
- Mean value theorem and extreme value theorem also do not hold


Explain l'Hospital's rule - maybe in a starred section

Regel von de L’Hospital verstehen – statt nur naiv anwenden!
https://www.youtube.com/watch?v=gV7TXoS_Bxc  
-Has some good counterexamples for when the rule doesn't work
 -actually, using it to establish that sin(x)/x should be 1 at x=0 involves circular reasoning
-At the end, gives a discrete analog (Stolzscher Grenzwertsatz)
 https://de.wikipedia.org/wiki/Satz_von_Stolz
 Maybe state that in the section about sequences

\end{comment}