\chapter{Combinatorics}
Combinatorics is concerned with counting all the possible ways to do a certain thing. This "thing" could be something like "select $k$ elements from a set of size $n$" or "create all possible acylic graphs of size $n$" or "find all ways to express a natural number $n$ as sum of smaller numbers" or "find all the possible ways that 5 rolls of a die could result in a sum less than 17". Things like that. The common theme is to count the number of different possibilities to do that thing. The result of such a counting tally will always be a discrete quantity or, more specifically, a natural number. Therefore, combinatorics is firmly situated in the realm of \emph{discrete mathematics} which is the mathematics that deals with things that can be enumerated, i.e. put into a (potentially infinitely long) list. The protypical example of a set of enumerable things is the set of natural numbers - other examples are integer numbers and the rational numbers\footnote{Yes, rational numbers are indeed enumerable - we'll see that later in axiomatic set theory.} but also things that are more common in computer science like trees, graphs, strings, etc. Combinatorics lays a lot of groundwork for other mathematical fields such as set theory, graph theory, number theory and probability theory. It is therefore appropriate to put the topic first in the "discrete math" part of the book. By the way, the examples of possible questions "select $k$ elements..." were taken from these fields - in that order. One typical feature of this number of possibilities to do a given thing is that it typically grows very fast with the problem size. In many combinatorial problem settings, we'll have to deal with huge numbers even for moderate problem sizes. For example, the number of ways to order 10 objects is given by 3628800 and for 20 objects, it's already 2432902008176640000. This is called combinatorial explosion. ...TBC...


% Mention combinatorial optimization. In priciple, it's possible to produce a solution to a combinatorial optimization problem by brute force, i.e. just trying all possibilities and select the best. But in practice, this is not possible because the numbers grow way too fast. Maybe mention tpyical problems from computer science: traveling salesman problem, boolean satisfiability, etc. I think, we typcially deal with at least exponential growth and often it's even superexponential.

% https://en.wikipedia.org/wiki/Combinatorial_explosion




% Other fields where combinatorics features prominently: computer science, statistical mechanics

% Discrete mathematics is the kind of mathematics that can represented exactly on a computer. To deal with continuous mathematics, we have to make do with approximations

% discrete: "how many", counting
% continuous: "how much", measuring

% ToDo: Mention combinatorial explosion. It refers to the fact that the number of ways to do the things that we are interested in, typically grows very fast.

\begin{comment}

https://en.wikipedia.org/wiki/Combinatorics
https://de.wikipedia.org/wiki/Abz%C3%A4hlende_Kombinatorik
https://en.wikipedia.org/wiki/Generating_function

Catalan, Stirling, Bernoulli ... numbers

Superpermutations: the maths problem solved by 4chan
https://www.youtube.com/watch?v=OZzIvl1tbPo
-shortest sequence of elements a,b,c,... that contains all possible permutation. For example,
 "aba" is a superpermtation of the elements a,b because it contains both possible permutations, namely
 "ab" and "ba" as substrings
 
Intro to Discrete Math - Welcome to the Course!
https://www.youtube.com/watch?v=rdXw7Ps9vxc&list=PLHXZ9OQGMqxersk8fUxiUMSIx0DBqsKZS 
by Trefor Bazett. Also treats logic and set theory

A surprising application of complex numbers in combinatorics #SoME2
https://www.youtube.com/watch?v=_lO6dja0CbA


The longest mathematical proof ever
https://www.youtube.com/watch?v=nfynJIb5tyg
-It's about Schur numbers. They are defined as: s(n) defined as: take the set of natural numbers 
 A = 1,2,3,...,n. What is the smallest number of colors k such that when giving each number one of 
 the colors, I can find triples a,b,c \in A that have all the same color and satisfy  a + b = c
-Is this also a combinatorics problem? More general, do problems of the kind: what is the smallest
 number such that..., what is the largest number such that...,i.e. the condition doesn't ask for an
 exact number but for a lower or upper bound for some condition? 
-What about "Ramsey theory" ...yes - it's combinatorics:  
 https://en.wikipedia.org/wiki/Ramsey_theory


https://en.wikipedia.org/wiki/Infinitary_combinatorics


The Dark Side of Pascal's Triangle #SoME4
https://www.youtube.com/watch?v=rQzu5JUjaG0
-About continuation of Pascal's triangle, Newton-Gregory formula, finite differences, discrete 
 calculus, combinatorial identities, umbral calculus

The Dark Side of Pascal's Triangle #SoME4
https://www.youtube.com/watch?v=xP8Z7IgIIZs
-Re-Upload?

Parity of permutations, impossible puzzles and the magical determinant
https://www.youtube.com/watch?v=rUiulWItECQ
-Parity can be calculated by considering the number of inversions. An inversion is when
 p(m) < p(n) and m > n  where m,n are the elements,indices in natural order and p(k) is the 
 permuted version of k. That means, before the permutation n comes before m and after the 
 permutation p(n) comes after p(m)


\end{comment}