\section{Differential Geometry}
In algebraic geometry, we considered zero sets of systems of polynomial equations. Under certain circumstances, these zero sets define a manifold and sometimes, we could find a parametric representation of such a manifold, i.e. a formula, that takes as input one or more independent parameters and spits out a point on the manifold as output. Differential geometry (DG) typically takes a parametric representation, i.e. a vector valued function of possibly multiple arguments, of such a manifold as a starting point. Since we will need to take work with (partial) derivatives of such a parametric representation, we will assume that all the functions are differentiable sufficiently often.

\medskip
DG can be roughly split into "classical" or "elementary" DG and "modern" DG. The former deals with curves in the 2D plane and with curves or surfaces in 3D space and is typically expressed in (a subset of) the language of vector calculus. The latter deals with the completely general case of $kD$ manifolds in $nD$ space where $k \leq n$ and is typically expressed in (a subset of) the language of differential forms (aka exterior calculus) and/or tensor calculus. Important applications of classical DG are in computer graphics, modern DG features prominently in general relativity.

\medskip
A very important, maybe the most important, concept in DG is the idea of curvature.

% curves vs surfaces vs general manifolds, intrisic vs extrinsic

...tbc...



\begin{comment}
	
In algebraic geometry, we deal with a manifold that is defined by an implicit equation of the form f(x,y) = 0 in 2D or f(x,y,z) = 0 in 3D or a system of such equations like f(x,y,z) = 0, g(x,y,z) = 0 in 3D. The equations are assumed to be polynomials but (I guess) the general ideas can be generalized to arbitrary equations. The manifold is typically of a dimensionality given by the dimensionality of the embedding space minus the number of equations (verify!). When I say "typically", I mean that there may be other cases but these can be considered as somehow being degenerate edge cases (verify!). For example, if we have have 1 equation in 3 variables, we normally get a 2D manifold, i.e. a surface in 3D space. 

In differential geometry, we also deal with manifolds but here, the manifolds are not defined via an implicit equation but rather by a set of parametric equations. We can imagine the parametrization to define a map of the manifold. We may have different parametrizations (i.e. maps or charts or patches) for different regions. The totality of all these maps (i.e. the set of all the maps) is called an atlas. Ideally, the maps should cover the whole manifold. Different maps may overlap with respect to the region that they cover. If they do overlap, they should agree in the regions of overlap. ...tbc...

In some sense, algebraic geometry is simpler than differential geometry because it just involves solving algebraic equations. Differential geometry, on the other hand, uses a lot of calculus which is usually considered to be a more advanced math topic than algebra. But in some other ways, differential geometry might also be considered easier that algebraic geometry because in DG we have the luxury of getting a parametrization of our manifold handed to us on a silver plate. Such a parametrization is for many purposes a much more convenient description of the manifold than the unwieldy implicit equation that AG gives us. One could actually say that the availability of the parametrization \emph{enables} us to make use of a lot of calculus that would unavailable otherwise.


https://mathworld.wolfram.com/Atlas.html
https://en.wikipedia.org/wiki/Atlas_(topology)#Charts

ToDo:
-There are at least 2 ways of defining a surface in 3D: 
 -parametrically as x(u,v), y(u,v), z(u,v) - that's the DG approach
 -implicitly as F(x,y,z) = 0 - that's the AG approach
-An explicit form z = f(x,y) can be seen a special case of both:
 parametric: x(u,v) = u, y(u,v) = v, z(u,v) = f(u,v)
 implicit:   F(x, y, f(x,y)) = 0
-Explain, how we can get a local parametrization from an implicit equation using
 partial derivatives...I'm not sure, if that's possible but I think, it should be.
 See:
 https://tutorial.math.lamar.edu/classes/calciii/parametricsurfaces.aspx
-Maybe we could also try to convert locally to an explicit form. It would become locally
 a 1D root-finding problem of F(x,y,z) = 0. Here, x,y are plugged in and can be seen as 
 parameters and we want to find z.
-Explain how we may get an implicit form from a given parametrization
-Sochi's book says that DG is more concerned with local features of the manifolds whereas
 AG is more about global features.
-I think, the fundamental local feature in DG is the metric tensor. Most other quantities of
 interest can be derived from it. In the case of surfaces, the metric tensor is just a 2x2 
 matrix. It can be computed from the extrinsic description of the surface via its 
 parametrization but one often just assumes it to be *somehow* given without caring much about 
 the how. It is an intrinsic feature of the manifold that can also be measured from an observer 
 who lives inside the manifold (Like A. Square living in flatland). Derived quantities are
 the different forms of curvature (except geodesic curvature which is an extrinsic property), 
 the Christoffel symbols, etc.
-Global features are things like the number of stationary points, orientability, presence or 
 absence of self-intersections, topological features such as the genus, ...

-Maybe use a circle of radius 5 as example curve and represent it 
 parametrically: c(t) = (5 cos(t), 5 sin(t))
 implicitly:  x^2 + y^2 - 25 = 0
 explicitly: y = +- sqrt(25 - x^2)
 and compute the tangent vector at a point (say (4,3)) via all 3 representations. Investigate, 
 how we can, from a given implicit equation and a point on the curve, compute coordinates of 
 another (infinitesimally) nearby point. Parametric and explicit representations can be used
 straightforwardly to "generate points" - but how can the implicit equation be used for that?
 1: by considering it a root-finding problem, 2: by assuming that we have a point given and
 somehow obtain a nearby point? ..which we may or may not refine using a root-finder?

%An Amazing Theorem for Tangents (Fabricius-Bjerre Theorem)
%https://www.youtube.com/watch?v=7_UclB2dB0o

Differential Geometry: The Intrinsic Point of View #SoME3
https://www.youtube.com/watch?v=jWoyBlIXVd8
-Gaussian curvature, being the product of two extrinsic quantities (namely, the two principal
 curvatures), turns out to be an intrinsic property. That's what Gauss found remarkable, according
 to this video. The "Theorema Egregium" itself is of a more quantitative nature, though.
-The Riemann curvature tensor is the nD generalization of Gaussian curvature.


https://en.wikipedia.org/wiki/Lie_derivative

How mathematicians describe curvature of surfaces
https://www.youtube.com/watch?v=UYiAlYlSwBo

\end{comment}