%\documentclass[12pt, twocolumn]{article}
\documentclass[12pt]{book}  % todo: use book and for the top-level items use "chapter"
%\usepackage{fullpage}           % makes all margins 1 inch?
\topmargin=-1.0cm
\textheight=23cm
\evensidemargin=-1.0cm
\oddsidemargin=-1.0cm
\textwidth=19cm
\setcounter{secnumdepth}{-1}  % suppress numbering of sections
\usepackage{amsmath}
\usepackage{amssymb}          % for mathbb
\usepackage{hyperref}

\usepackage{comment}          
% to comment out larger sections via \begin{comment} ... \end{comment} 
% see:
% https://tex.stackexchange.com/questions/17816/commenting-out-large-sections
% https://tex.stackexchange.com/questions/11177/how-to-write-hidden-notes-in-a-latex-file/73418

%\DeclareMathOperator{\d}{d}                  % exterior derivative

\begin{document}
	
% formatting:
\parindent=0in
\parskip=0pt
\pagenumbering{roman}
		
% main text
\pagenumbering{arabic} \setcounter{page}{1}

\author{Robin Schmidt}

\title{Mathematical Recipes for Scientists, Engineers and Programmers}
\maketitle
\tableofcontents

\chapter{Basics}
\include{Basics_Intro} 
\include{Basics_Foundations} 
\include{Basics_Numbers}
\include{Basics_Functions}

\chapter{Linear Algebra}
\include{LinAlg_Intro} 

\chapter{Calculus}
\include{Calc_Intro} 
\include{Calc_LimitsAndContinuity} 
\include{Calc_Differentiation}
\include{Calc_Integration} 
%\include{Differential_Equations}
%\subsection{Series}
%\subsection{Integral Transforms} % Fourier- and Laplace trafo
%\subsection{Integral Equations}
%ToDo: finite differences, summation calculus, difference equations, z-trafo
% see https://www.youtube.com/watch?v=4AuV93LOPcE

%\section{Analytic Geometry}
% maybe this fits best directly after Linear Algebra as application thereof, so maybe place Calculus before Linear Algebra but maybe if we also compute areas of shapes via integrals, calculus needs to be introduced before geometry as well...maybe be a bit more general and include some (basic) calculusly stuff like areas and volumes, optimization of shapes (cylindrical can with smallest surface area kind of stuff)..no multivariable calculus, though - that's for later. but maybe introduce bivectors and trivectors already - they are simple and intuitive and will be needed for exterior algebra


\chapter{Multilinear Algebra}
Multilinear algebra is the study of operations that take multiple inputs and are linear in each of those inputs. For example, consider the sandwich product of two vectors and a matrix: $\mathbf{z = x^T A y}$. The matrix $\mathbf{A}$ can be thought of defining an operation that takes two vectors as inputs and produces a scalar as output. The so defined operation is indeed linear in both inputs in the sense that it makes no difference whether you add or scale at the input or output side.
 % covectors/k-forms, tensors, ..

\chapter{Multivariable Calculus}
Multivariable calculus is the calculus of continuous functions that may depend on several input variables and/or produce several output variables. As with single variable calculus, one may find minima and maxima of a function with multiple inputs using derivatives by requiring them to be zero (yes: derivatives in plural!). This is important in optimization. Because we now have several input variables, we may take derivatives with respect to any of them which is why they are called partial derivatives in this context. We can collect them in a vector which we then call the gradient. The tool to compute areas under functions - the integral - generalizes to compute other aggregate quantities such as lengths of curves and volumes and surface areas of 3D objects, the amount of a fluid flowing through a surface, the work done by a force field on an object moving through that field, etc. The multivariable version of a differential equation is a partial differential equation, named like that because it involves partial derivatives. So, despite its name, a "partial" differential equation is actually a (much) more complicated thing than an ordinary differential equation. A subfield of multivariable calculus, called vector calculus, is mainly concerned with functions that live in the same 3-dimensional space that we do, making it suitable to describe many physical phenomena. The 3D vector calculus uses some special notations that are specific to 3D and do not readily generalize to an arbitrary number of dimensions. Most notably among them is the curl which is defined by a certain cross-product which itself makes only sense in 3D. Sometimes the equations are boiled down to just two dimensions, which you can easily embed in 3 dimensions, so with some hacks (the curl becomes a scalar) it kinda works there, too. Sometimes the equations also involve time which is then treated seperately from the spatial dimensions. Time being treated as "something else" than space coincides with our daily experience but makes it less than ideal in the context of relativity where time is treated on equal footing, which is why you often see different formalisms used there. Alternatives to vector calculus that do generalize to higher dimensional spaces are exterior calculus (a.k.a. calculus of differential forms), tensor calculus or geometric calculus. All these calculi are based on their respective algebras each of which defines a particular product and the problematic cross-product from 3D vector algebra is not a thing anymore in these algebras.
%\subsection{Basics} \include{MultivariateCalculusBasics} 
% multivariate functions, partial derivatives & gradient, Hessian, Jacobian, directional derivative, multiple integrals, (constrained) optimization (also numerical), parametrized curves/surfaces

%\subsection{Vector Calculus} \include{VectorCalculus} 
% scalar and vector fields, curl, div, line/surface/volume integrals, Green/Gauss/Stokes theorem,

\section{Exterior Calculus} 
\include{MulCalc_ExteriorCalculus} 

%\subsection{Partial Differential Equations}
% transport, heat, wave, laplace/poisson, schrödinger, maxwell, navier-stokes, Einstein, Dirac
%give them all in all possible formalisms (vector, exterior, tensor, geometric)...or maybe make a separate chapter "Phyiscs", give Maxwell's equations in the following formalisms: (1) original form in terms of partial derivatives, (2) vector calculus (div, curl, etc.), (3) integral form, (4) differential forms, (5) tensors, (6) geometric calculus

% sections: vector calculus, exterior calculus, geometric calculus, tensor calculus (ascending generality)
% scalar fields, vector fields, partial differentiation (grad, Jacobian, Hessian, Laplacian, divergence, curl), integral theorems (Green, Gauss, Stokes), differential forms, generalized Stokes theorem, Tensor fields


\section{Complex Calculus}
Complex calculus is the calculus of functions that map a complex number to another complex number. Since input and output are both 2-dimensional, it makes sense to look at complex calculus from the perspective of 2D vector fields. For a function $w = f(z) = u(x,y) + \i v(x,y)$ to be differentiable in the complex sense, the partial derivatives of the component functions $u,v$ with respect to $x,y$ must exist and satisfy the Cauchy-Riemann (CR) differential equations: $u_x = v_y, u_y = -v_x$. This condition has wide ranging implications which allow a lot of simplifications to be made in the computations of path integrals. This can also help to evaluate certain definite real integrals that would otherwise be hard to do. If the function $f$ is considered as a mapping that maps points from the $z$-plane to the $w$-plane, the CR equations imply that the images of curves in the $z-$plane that intersect at a given angle, will intersect at the same angle in the $w$-plane. Such an angle-preserving mapping is also called a conformal mapping. If a function is complex differentiable at a given point $z_0$, this implies that all higher order derivatives also exist and that the function can be locally expanded in a Taylor series that converges in a circular region whose radius extends to the nearest critical point [verify!]. A critical point is...
%todo: say something about multivariable complex calculus

%\section{Approximation}
%\paragraph{Trigonometric Interpolation} We have a bunch of datapoints $x_k, y_k$ and want to find 



\begin{thebibliography}{10}
\bibitem{MatheRezepte}
Christian Karpfinger. \newblock {\em Hoehere Mathematik in Rezepten, 3. Auflage}.
\end{thebibliography}	

\bibliography{Bibilography} 
\bibliographystyle{plain}  

\begin{comment}
-include python code in the text, realizing numerical and symbolic computations using numpy and sympy
 -there should be a library accomanying the book

\end{comment} 
	
\end{document}


