\chapter{Misc Math Topics}

\section{Beautiful Formulas}
Mathematics is not just super useful in science and engineering but also very beautiful in it own right. In fact, the practical and useful aspects of it are sometimes sneered upon by (pure) mathematicians. They do it not for its practical applications but purely for their intellectual enjoyment of their ivory tower level mind blows. ...TBC...

% What makes an formula beautiful? It should be profound, i.e. have huge implications, give great inseight. Maybe it should b simple to state but difficult/non-obvious to derive? 

\paragraph{Euler's Formula}
The following formula:
\begin{equation}
\e^{\i \pi} + 1 = 0
\end{equation}
has been called by many mathematicians the most beautiful formula of all of mathematics because it establishes a connection between the 5 most important mathematical constants $0,1,\i,\pi,\e$. Furthermore, it makes use of each of the 3 important mathematical operations addition, multiplication, exponentiation...TBC...

% Special case of
% 0,1,i: algebra, pi: geometry, e: analysis/calculus

% algebra, geometry, calculus

\paragraph{Euler Product}
For a given complex valued parameter $s$ with real part greater than one the following equation holds:
\begin{equation}
\sum_{n=1}^{\infty} \frac{1}{n^s} = 
\prod_{p \in \mathbb{P}} \frac{1}{1 - p^{-s}}
\end{equation}
The real part must be greater than one in order to have convergence of the sum on the left hand side. I think the product on the right hand side converges in a namely domain, namely for $\Re(s) > 0$ [VERIFY!]. The formula establishes a relation between a sum over \emph{all} natural numbers with a product over only the \emph{prime} numbers. The sequence of all (positive) natural numbers $(1,2,3,4,\ldots)$ is very simple and well understood but the sequence of prime numbers $(2,3,5,7,11,13,17\ldots)$ is very mysterious. Therefore, establishing a connection between a very simple sequence of numbers and a more complicated one may help to gain insights into the more complicated sequence. We can also interpret both expressions as defining a function in the domain where they converge. This function of a complex variable $s$ is an analytic function and known as the Riemann zeta function and denoted by $\zeta(s)$. In those parts of the complex domain where neither se sum nor the product converges, the function can be defined via analytic continuation. Of special interest are the zeros of the function...TBC...

% Number theory

\paragraph{Generalized Stokes Theorem}
Given a manifold $M$ and a differential form $\omega$ defined on the manifold with exterior derivative $d \omega$, the following relation holds:
\begin{equation}
\int_{M} d \omega = \int_{\partial M} \omega
\end{equation}
The left hand side is an integral over the whole manifold and the integrand is the exterior derivative of the differential form. The right hand side is an integral over the boundary of the manifold and the integrand is the differential form $\omega$ itself. This equation is a vast generalization of all the integral theorems of vector calculus. It also includes the fundamental theorem of calculus as a very simple special case. Without much exaggeration, one could say that it encapsulates the essence of calculus. ...TBC...

% Calculus

% I think, it also encompasses soem integral formulas from complex analysis

\paragraph{Fundamental Theorem of Geometric Calculus}
The generalized Stokes theorem can be generalized even further when broadening our perspective from exterior to geometric calculus. ...TBC...

% We now must consider the geometric derivative...


\begin{comment}

The most beautiful equation in math.
https://www.youtube.com/watch?v=u7BadbVwnl4&lc=Ugxn_M9K5WiVds15A-Z4AaABAg
...has a comment on what all these equations are:
from left to right, top to bottom: Euler's identity, Navier–Stokes equation, (generalized) Stokes's theorem, Riemann–Roch theorem, Euler's factorization of the Riemann zeta function, Gauss–Bonnet theorem, ???, Hardy–Ramanujan–Rademacher partition formula

Try to find the most beautiful formulas for different ares of math:
-Abstract algebra: Lagrange's theorem (group theory), Fundamental theorem of Glaois Theory?
-Algebraic geometry: Bezout's theorem,
-Set theory: Continuum Hypothesis,
-Statistics: Central limit theorem / Gaussian distribution
-Calculus of Variations: Euler-Lagrange equation
-Tensor calculus:
-Linear algebra: SVD? rank-nullity theorem? Something about eigenvalues? the fact that a
 matrix is a zero of its own characteristic polynomial?
-Differential geometry:
-Topology:
-Functional analysis:
-functional equation for exp or log?



\end{comment}