\chapter{Philosophical Principles}
This chapter is a bit less mathematical and more philosophical and maybe meta-mathematical. We look at some general underlying principles of mathematics ...TBC...

% Meta-math


%###################################################################################################
\section{Elements of Math}

\subsection{Definitions}


\subsection{Theorems}

\subsection{Proofs}


\subsection{Conjectures}
A conjecture is a proposition that is seen as a potential candidate for a possible theorem. But the truth of the proposition has not yet been proven. Conjectures are typically formed when one plays around with some problem and discovers some apparent regularities, rules or laws which seem to be true but are not yet known facts.  ...TBC...

\subsubsection{Empirical Evidence}
To come up with conjectures, one can take the tried and true approach of the empirical sciences which is called the \emph{scientific method}. It consists of a cycle of (1) discovery of patterns, (2) attempt of explanation, (3) making predictions based on the explanation attempt, (4) design of experiments to verify the predictions, (5) execution of the experiments, (6) examination of the experimental results - back to (1). In mathematics, the "experiments" will typically involve a computer program. For example, it's possible to write a program that checks a proposition like the Goldbach conjecture experimentally up to some large number $N$. If the conjecture fails for some $n \leq N$, the conjecture has been falsified - it's now known to be false. If it holds up up to $N$, we may (very tacitly!) treat that as evidence, that it might hold up in general. This evidence in mathematics is sometimes called \emph{numerical evidence}. In contrast to the natural sciences, empirical evidence - however overwhelmingly abundant it may be - is never a substitute for a proper proof to establish the truth of a proposition. Modern mathematics has stricter requirements and higher standards - it demands an airtight, rigorous proof.

...TBC...

% This is mainly because in mathematics, we tend to make propositions about infinite sets and the vastness of infinity is always (infinitely!) larger than any amount of finite evidence could ever be. However large the pile of empirical evidence may be, it is still finite and therefore basically nothing compared with infinity. That statement can even be made a bit more mathematical: The ratio of any finite number $N$ to infinity is always exactly zero: $\frac{N}{\infty} = 0 \forall N \in \mathbb{N}$. That's true for the humble $N = 100$ just the same as it is for a giant googolplex of $N = 10^{10^{100}}$ or Graham's number or whatever wild googology-construction anyone may ever come up with. Now - while this is certainly fromally true, in practice, of course, having verified some conjecture about natural numbers up to $N = 10^20$ would still count as stronger evidence than having it verfied only up to $N = 100$.

% ToDo: Bring examples from number theory where some conjecture seemed to hold up up to really large numbers but was nevertheless proven to be false.


%###################################################################################################
\section{Abstraction and Generalization}
Mathematics in infamous for being very abstract. But was does "abstract" mean? Abstraction in mathematics is the process of recognizing, identifying and explaining common patterns in a bunch of concrete problems and then trying to find a solution for all these concrete problems at once by finding a solution to the abstract, generalized version of these problems. Solving a problem of the same type later again will then be much easier - we will just need to recognize that it again a concrete instance of an abstract problem that we have already solved. We can then just apply our general solution. ...TODO: explain generalization, permanence principle

% Generalization, Permanence, Pattern recognition/identification/explanation


%###################################################################################################
\section{Reductionism and Emergence}

% Complexity from simplicity - deconstruction to simpler parts...there's a name for that
% emergence? Reductionism - I think, that was the word
% https://en.wikipedia.org/wiki/Reductionism
% -relevant also in the natural sciences

% - Reductionism: Try to break down a complex thing into simpler constituents, understand the 
%   behavior of the individual constituents
% - Emergence: The behavior of the bigger system then "emerges" from the collective behavior of
%   constituents.
%
% - Applied to math: When we try to understand what numbers are, we may try to "model" their 
%   behavior with simpler (or at least more fundamental) objects. This is done when expressing 
%   numbers in terms of sets. I personally do not think that this set theoretical explanation is
%   really what numbers really "are" - it's just a way to "model" their behavior with more 
%   fundamental objects. If these objects are "simpler" may be debatable. Their desciption may not
%   be. This is also the case, for example, for the beahvior of a tennis ball and the
%   quantum-mechanics of the molecules that make up the ball.

%###################################################################################################
\section{Duality}
An interesting observation across many fields of mathematics is that we sometimes find objects that in some sense are a sort of "mirror image" of other objects. ...TBC...examples: vectors/covectors, vectors/pseudovectors, points/lines, dual graphs, dual platonic solids, ...

% schlüssel/schloss prinzip
% Symmetry, Isomorphism

% What about inversion as in differentiate/integrate
% adjoint matrices?

% https://en.wikipedia.org/wiki/Duality_(mathematics)
% https://en.wikipedia.org/wiki/Dual_lattice
% https://en.wikipedia.org/wiki/Duality_(order_theory)

%
% Dual Spaces: The 'Underworlds' of Mathematics
% https://www.youtube.com/watch?v=rpBAUxXI1kY
% -Has an interesing notational idea: denote an elemnt of the dual space with the multiplication 
%  dot after it, i.e. dual(a) = a \cdot

%###################################################################################################
\section{Unification}
The more math you learn, the more unexpected connections and bridges between seemingly unrelated fields of math become apparent. It seems like somehow everything hangs together in a vast interconnected network of mathematical knowledge. 

%as if there is a greater underlying unified theory of all (or at least much) of mathematics. ...TBC...

% example: Langlands program


%###################################################################################################
\section{The Axiomatic Method}

% See: ABoST - explained "undefined concepts" by examples like points, lines, parallellism, etc.
% in geometry.
% -Give examples: Euclid's axioms for geometry, Hilbert's axioms for geometry, Peano's axioms for
%  arithmetic of natural numbers, Zermelo-Frenkel's axioms for set theory, von Neumann's axioms for
%  set theory, axiomatically postulating the real numbers (as complete ordered set)


\begin{comment}

-Formalism vs Intution
-Precision, Non-Ambiguity of langauge
-Automation/mechanization of thought (formalism, algorithmic proof checkers)
-Finitism, Constructivism

https://en.wikipedia.org/wiki/Constructivism_(philosophy_of_mathematics)

Unexpected connections:

The Trig Hiding Inside the Factorials (And Harmonic Numbers)
https://www.youtube.com/watch?v=-HwwY4czx_E

Is Math Discovered or Invented
https://www.youtube.com/watch?v=jsyalJgzXT4

Restarting Mathematics With These 5 Fields
https://www.youtube.com/watch?v=AJ9ddGejUcM

-Langlands program

\end{comment}