\chapter{Topology}
Topology can be described as the study of those properties of geometric objects that are preserved under continuous deformations of the object. An example of such a property is the number of holes that it has: a sphere has none, a torus has one. There are also shapes that have two, three, etc.. That number is such a topological property. You may continuously deform the object as much as you like - the number of holes will remain the same. Intuitively, for such a deformation to qualify as "continuous", it shall not rip or cut apart material, poke holes into it, glue things together that formerly were apart, close holes, etc. Mathematically, the class of allowed deformations is defined in terms of \emph{a topology} which is a specific set of subsets of a set $S$. Note how the term "topology" is overloaded as referring to two things: to the mathematical field itself and to a particular mathematical definition of a set of a certain kind that is used within that field. Such a topology will allow us to define continuity in a way that doesn't rely on a distance measure. Central to the field of topology are homeomorphisms which are invertible functions between two topological spaces which are continuous in both directions. In this context, the "topological spaces" are interpreted as geometric shapes and the function transforms one shape into another.

% Topology revolves around the notion of "closeness". When are two elements close to each other? What are th maps that preserve closeness?

%such as the surface of

%homotopies, which are continuous transformations of one function into 

%Topology has been called "rubbersheet geometry" by some authors which captures the idea quite nicely. ...TBC...

% Explain also the term "point set topology". Topology is concerned with sets of points.

\begin{comment}

% knot theory, moebius strips

https://en.wikipedia.org/wiki/Topology
https://en.wikipedia.org/wiki/Topological_space

https://en.wikipedia.org/wiki/Topological_property

https://en.wikipedia.org/wiki/Homeomorphism
https://en.wikipedia.org/wiki/Homotopy

Grundbegriffe der Topologie: offene Mengen und Umgebungen
https://www.youtube.com/watch?v=55jUgqwFnec&list=PLb0zKSynM2PD3i3xMuWrUF9_txMrJMGEZ&index=23
in the playlist about differential geometry


https://de.wikipedia.org/wiki/Satz_von_Stokes
https://de.wikipedia.org/wiki/De-Rham-Kohomologie#Satz_von_de_Rham


What is a hole?
https://www.youtube.com/watch?v=IDcw33YRgpY
-about algebraic topology
-sequel of "What is algebraic geometry?" https://www.youtube.com/watch?v=MflpyJwhMhQ
-A space has "no holes" iff every loop in it is homotopic to (i.e. can be contracted to) a point
-one loop is homotopic to another, if it can be continuosuly be transformed into it

Examples of topological invariants (verify!):
-number of holes (sphere vs torus)
-number of sides (loop vs Moebius strip)
-minimal number of crossings in a knot
-winding numbers (of a given path with respect to a given point/hole?)?
-dimensionality of an objcet and/or hole?


https://www.youtube.com/watch?v=unikWDPBY8M  Topology is Impossible Without These 7 Things

Maybe graph theory can be thought of as "discrete topology"? Both fields deal with connectivity between members of a set. Both have a concept of paths - and the meaning of these concepts is compatible. Maybe topology could be thought of as continuous graph theory where there is a continuum of nodes? How could a path deformation look like on a discrete graph? Maybe replace one node with its two adjacent edges at a time? How could we formalize the idea of "adjacent paths"? Maybe a path P and a path Q could be defined as being adjacent, when Q can be created from P, by replacing one node in P by some subpath? when we have 3 connected nodes within a path a--b--c, we cut out --b-- and replace it by some other subpath that connects a and c like a--x--y--c. I think, we need to allow the replacement path to be a subpath of arbitrary length

https://en.wikipedia.org/wiki/Algebraic_topology

Topology vs "a" Topology | Infinite Series
https://www.youtube.com/watch?v=tdOaMOcxY7U&list=PLa6IE8XPP_glwNKmFfl2tEL0b7E9D0WRr&index=16


The Insane Math Of Knot Theory
https://www.youtube.com/watch?v=8DBhTXM_Br4

https://en.wikipedia.org/wiki/Trefoil_knot

Knots, Links, Braids


How Did Mathematicians Abstract the Idea of Distance?
https://www.youtube.com/watch?v=UDP7nDV_734

The Concept So Much of Modern Math is Built On | Compactness
https://www.youtube.com/watch?v=td7Nz9ATyWY

Topological Spaces Visually Explained
https://www.youtube.com/watch?v=B-Y3-XpAdMU
- inner product space < normed vector space < metric space < topological space
- trivial aka indiscrete topology: { X, {} }
  discrete topology: set of all subsets
  -> these are the extreme cases and the interesting topologies fall somewhere in between
- a topology is the weakest structure that you can establish on a set that allows these two
  notions: (1) convergence of points, (2) continuity of maps
See also:
- https://oeis.org/A000798  1, 1, 4, 29, 355, 6942, 209527, 9535241, 642779354, ...
  Number of different quasi-orders (or topologies, or transitive digraphs) with n labeled elements.
- https://en.wikipedia.org/wiki/Preorder
- https://arxiv.org/abs/1503.08359  On The Number Of Topologies On A Finite Set


A Sphere is a Loop of Loops (Visualizing Homotopy Groups)   by Dennis Sweeney
https://www.youtube.com/watch?v=CxGtAuJdjYI

A Quick Intro to Fiber Bundles (Hopf Fibration)
https://www.youtube.com/watch?v=dkyvZo68IoM

The Topology of the Cantor Set
https://www.youtube.com/watch?v=c4oKdtUtJOk

The fascinating link between Topology and Politics: an introduction to TDA.
https://www.youtube.com/watch?v=KhHDfbwALX8

Associahedra: The Shapes of Multiplication | Infinite Series
https://www.youtube.com/watch?v=N7wNWQ4aTLQ
-Loop multiplication: Let a = a(s), b = b(s) be two loops, i.e. two functions from [0,1] into a
 topological space that define closed curves with a(0) = a(1), b(0) = b(1). Let the strat point of b
 also coincide with the end point of a: b(0) = a(1) such that it makes geometric sense to
 concatenate them. Then the "multiplication" of these two loops is defined as concatenation of a and
 b (traversed at twice the speed): 

   (a b)(s) = a(2s) for 0 <= s <= 1/2  and   b(2s-1) for 1/2 <= s <= 1

 I think, the definition of the product as concatenation can make sense for any pair of curves
 for which a(1) = b(0). Maybe even open ones. But: If we have only this condition and not also
 b(0) = a(1), we could only form the product ab but not ba, I guess?
-This mulitplication is *not* associative because the location to which a particular time instant s
 is mapped is different in (ab)c and a(bc). But the two results can be stretched to one another by a
 homotopy, namely
 H(s,t) = ...
-More on the topic:
 https://www.math3ma.com/blog/multiplying-non-numbers
 https://www.claymath.org/library/academy/LectureNotes05/Lodaypaper.pdf
 https://arxiv.org/pdf/math/0212126

Turn a number system into a geometric space. Intro to algebraic geometry, Spec(R) and schemes.#SoME4
https://www.youtube.com/watch?v=6q9LipVbI80
-Says at 1:30 that "a topology" is a "notion of nearness"

Topological spaces explained
https://www.youtube.com/watch?v=9hbYNLWb8BY

The Most Obvious Theorem in All of Mathematics
https://www.youtube.com/watch?v=pLgcZLysOFk
-A video on the Alexander horned sphere as a counterexample to the Jordan-Schoenflies theorem
 generalizing to three dimensions. 


https://en.wikipedia.org/wiki/Metric_space#The_topology_of_a_metric_space

A Beautiful Math Problem with an Incredible Solution
https://www.youtube.com/watch?v=O_HqzoNYXks

\end{comment}